TRITON: Globally Consistent Sim-to-Real Transfer

### 2022-06-12 CORL/example.tex
Updates from Overleaf

%TODO:
%	Formulation Section
%	Figures
%		Main diagram - break it up into pieces with just text
%		UVL three-axis diagram
%		Replace C with (pi,s)
%	Method:
%		Unprojection Consistency Loss
%		Texture Reality Loss

%TODO:
%	Formulation Section
%	Figures: Main diagram - break it up into pieces with just text
%	Figures: UVL three-axis diagram
%	Method:	Unprojection Consistency Loss
%	Method:	Texture Reality Loss

\newcommand{\Tab}{T_{A\rightarrow B}}
\newcommand{\Tba}{T_{B\rightarrow A}}
\newcommand{\pih}{\hat{\pi}}
\newcommand{\sh}{\hat{s}}
\newcommand{\ph}{\hat{p}}
\newcommand{\pis}{\pi,s} % pi and s pairs from domain A
\newcommand{\pish}{\pih,\sh} % pi and s pairs from domain A
\newcommand{\ppis}{(\pis)} %Parenthesized pis
\newcommand{\ppish}{(\pish)} %Parenthesized pis

\newcommand{\toB}{_{1\dots B}}
\newcommand{\toBL}{_{1\dots B,1\dots L}}
\newcommand{\toL}{_{1\dots L}}

\newcommand{\lUC}{\ell_{UC}}
\newcommand{\lTR}{\ell_{TR}}

%%%%%%%%%%%%BIG TODO::::
%%%%%% MAKE BATCH SIZE N INSTEAD OF B - B is a domain!

@@

% \usepackage{corl_2022} % Use this for the initial submission.
\usepackage[final]{corl_2022} % Uncomment for the camera-ready ``final'' version.
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsfonts}

@@

% \title{TRITON: Globally Consistent Sim-to-Real Transfer
\title{Neural Textures make Better Simulators

``I can't believe it's not real!" %Definitely gonna delete this lol
 %via Neural Textures
 }

@@

	QUESTIONS FOR MICHAEL:
	\begin{itemize}
		\item - TODO: WE MAINTAIN WHAT WE NOW CALL "SURFACE CONSISTENCY" instead of view consistency! Or texture consistency? Etc? What should we call it?
	\end{itemize}

@@

TODO: Note that the simple input of UVL, or an RGB neural texture, means any simulator should be plug-and-play with this algorithm.

@@

\section{Formulation}
\label{sec:data}

TODO: Describe what UVL is ; 
Formulate the setting: What our inputs are and the targets are and what we learn (a neural texture function) and how we can take advantage of it for robotic learning.
Don't go into details; just say 'its a black box function'

TODO: Add UV to the fourier feature diagram.

TODO: Create subfigures with subcaptions. 
2a: Nothing but UVL=UV+L
2b. 3d uvl axis to texture pack (drew on whiteboard)
2c. Skip over the translation network and show a pretty robot arm picture

This is the formulatoin section; not data section. Dont talk about how many photos we use etc.

JUST IN CASE THINGS ARE ERASED:
We learn neural texture corresponding to each object, where each texture is represented as a function mapping... 
The idea is to learn object textures as a set of continuous functions from a set of (unlabeled) real images, and use them to generate realistic images.
Given a 3D structure of the scene represented with a UVL image, the learned neural textures will translate them into a realistic image.
When used for the robot learning, this enables learnable, realistic sim2real of robot training images.

IN THE INTRODUCTION: How we're using a neural neural network network texture. Also in related works.

TODO: Replace neural texture in figure 3 with axis, and treat it as a function that takes UV in with an arrow pointing to it.

BREAK IT DOWN APPROACH:
Don't have any pictures. Have breakdown where we add new things in multiple figures.
First figure: Just image translation. 
Seconf:
Third:
Etc...

@@

@@

\subsection{Neural Texture and Projection}

@@

	TRITON's neural texture module is what lets it learn the albedo of objects, and thus take advantage of 3d geometric priors. By learning this texture, we can encode features directly onto the surfaces of the 3d objects in a scene, which the image translation module uses to maintain visual consistency between states and viewpoints.

	Given a batch of $B$ UVL Scenes $s_{1\dots B}$ with $L$ different label values, we project $L$ different neural textures $\tau_{1\dots L}$ onto that scene; obtaining projections $\tau_{1\dots B}$. This projection mechanism is illustrated in figure \ref{fig:uvl_explanation}.

@@

	Each neural texture $\tau_i$ is a function consisting of two components: a multi-layer perceptron $M_i$ and a static set of 256 random spatial frequencies $f_{1...256}$. 
	Given a two-dimensional $uv$ vector, the spatial frequencies are used to generate a high dimensional embedding used as input for the MLP.
	\begin{equation}
			\tau_i(uv) = 
			\
			M_i\left(\left[
				\sin f_1 uv, \ 
				\cos f_1 uv, \ 
				\sin f_2 uv, \ 
				\cos f_2 uv, \ 
				\dots \ 
				\sin f_{256} uv, \ 
				\cos f_{256} uv \ 
			\right]\right)
	\end{equation}
	where $uv$ and $f_i \in N_{ormal}(\mu=0,\ \sigma=10)$ are two dimensional vectors,
	 $f_i uv$ represents their dot product, 
	%  $rgb$ represents a 3d color vector,
	 and $M_i$ maps a $uv$ embedding to an $rgb$ color value:
	 $M_i: \mathbb{R}^{256} \rightarrow \mathbb{R}^3$.
	Note that there is nothing special about 256 or 10: these hyperparameter values just happen to work nicely.

@@

\subsection{Image Translation}
\label{sec:image_translation}

% TO PUT IN FORMULATION:
% The simulation domain A (UVL maps concatenated to their texture projections) has 6 channels, whereas the real image domain B (photographs) has 3 channels. An element of A is $\ppis \in A$ and $p \in B$.
% $p \in B$ is a batch of real photos,
% and $\ppis\in A$ is a batch of simulated projection/uvl scene concatenations.

% TODO: Why doesn't citep work? The references don't show the authors names...it used to work...did I mess up example.bib?

% TODO: Get rid of the variable "C" and replace it with $\ppis$

% TODO: Is l2 loss capitalized or lower case?

Our image translation module, $T$, is based on the image translation module used in \citep{surgical_video_translation}, which is a based on a variant \cite{surgical_image_translation} of MUNIT \cite{munit}. The main differences between our image translation model and MUNIT are that our style code is fixed making it unimodal, noise is injected into the latent codes during training to prevent overfitting, and the reconstruction criterion we use. During inference however, this intermediate noise is removed and our image translation module is completely determinstic. Apart from these differences, our image translation module has the same loss formulations and generator/encoder/discriminator network architectures as MUNIT.

$T$ is a bidirectional image translation module, and has two translation directions. $\Tab$ translates sim to real (aka domain A to domain B), and $\Tba$ translates real to sim (aka domain B to domain A).

MUNIT translates images by encoding both image domains into a common latent space, then decoding them into their respective domains.
In our translation module, we define fake photos $\ph = \Tab(\pis) = G_B(E_A(\pis))$ and fake projection/uvl scenes $\ppish = \Tba(p) = G_A(E_B(p))$
	where $E_A, E_B, G_A, G_B$ are encoders and generators for domains A and B respectively.
% We also define the outputs of these translators as fake photos $\hat{p} = T_{A\rightarrow B}(\pis)$ and fake projection/uvl scenes $(\hat{\pi},\hat{s}) = T_{A\rightarrow B}(\pis)$.

% TODO: Choose one of the three equation representations. Which is best? The next paragraph will look less ugly when the equations are shorter.

To ensure cycle consistency, we have an image similarity loss $\ell_{cyc}$ to ensure
% $c \approx G_A(E_B(G_B(E_A(\pis)))) = T_{B\rightarrow A}(T_{A\rightarrow B}(\pis)) = T_{B\rightarrow A}(\hat{p})$
$\ppis \approx T_{B\rightarrow A}(\hat{p})$
and vice versa
% $p \approx G_B(E_A(G_A(E_B(p)))) = T_{A\rightarrow B}(T_{B\rightarrow A}(\pis)) = T_{A\rightarrow B}(\hat{c}) = T_{A\rightarrow B}(\hat{\pi},\hat{s})$. 
$p \approx T_{A\rightarrow B}(\hat{\pi},\hat{s})$.
%I only want to use one of the two representations; either using T_AB or GA(EB) etc. Which one should I use?
We also impose an image similarity loss $\ell_{rec}$ to ensure that $\ppis \approx G_A(E_A\ppis)$ and $p \approx G_B(E_B(p))$, which is effectively an autoencoder loss. 
In addition, we also have content similarity loss $ \ell_{con}$ to make 
$E_A(\pis) \approx E_B(\ph)$ and $E_A(p) \approx E_A(\pish)$.
% $E_A(\pis) \approx E_B(\Tab(\pis))$ and $E_A(p) \approx E_A(\Tba(p))$.

Our similarity losses are the sum of L2 distance and negative multi scale structural image similarity (MS-SSIM), introduced in \citep{msssim}. For our purposes, MS-SSIM is a metric that measures the similarity of two BCHW tensors on a scale from 0 to 1. Given two tensors $x$ and $y$, their similarity loss is $\mu\left((x-y)^2\right) - msssim(x,y)$.

We also have adversarial losses $\ell_{adv}$ that come from two discriminators $D_A$ and $D_B$, targeting domains A and B respectively. They use the LS-GAN loss introduced in \citep{lsgan}.

In total, the loss for the image translator module $T$ is $\ell_T=\ell_{cyc}+\ell_{rec}+\ell_{con}+\ell_{adv}$.

% NOTE: THe other paper \citep{surgical_video_translation} didn't do into this much detail; instead putting most of these losses in teh appendix. I think we can cut this down? I'll check with michael first...right now I'll include everything.

\subsection{Unprojection Consistency Loss}
\label{sec:unprojection_consistency_loss}

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=400pt]{../images/unprojection_consistency_loss.pdf}
	\end{center}
	\caption{
		This is the unprojection consistency loss used in figure \ref{fig:main_diagram}.
		The Unprojection Mean and Reprojection nodes are not used in any losses, and help to illustrate the next section \ref{sec:texture_realism_loss}.
		% After calculating fake photos $\hat{p}_{1...B}$ from UVL scenes $s_{1...B}$ as seen in figure \ref{fig:main_diagram}, 
		% we can unproject fake photos $\hat{p}_{1...B}$ back into texture space to obtain $u_{1...B,\ 1...L}$ .
		% If the fake photos display consistent textures along their 3d surfaces, then the unprojected textures should be similar.
		% We can measure this similarity by taking the per-channel standard deviation of these unprojections, aka $\sigma(u)$, then averaging the result.
		% When the fake photos display similar surfaces, the unprojections will be similar making the unprojection consistency loss low.
		% Additionally, we can also calculate the average unprojection, which is useful for visualization purposes (it will look blurry if $\hat{p}$ is inconsistent).
	}
	\label{fig:unprojection_consistency_loss}
\end{figure}

To maintain surface consistency, we sample a random batch of UVL scenes $S\toB$, and translate them into $\ph\toB$.
Using geometric knowledge of the scene encoded in $S\toB$, we individually unproject each $\ph_i \in \ph\toB$ back into texture space, to obtain $B$ sets of $L$ unprojected textures $u\toBL$.
We then use the per-pixel, batch-wise standard deviation of the unprojections as unprojection consistency loss $\lUC = \sigma(u\toBL)\toL$.
This loss observes how different these unprojections are from one another, and thus how different the object surfaces appear between translations.

Intuitively, if $\lUC$ were $0$, it would mean the object surfaces in translations $\ph$ appear exactly the same in every scene.

This loss is a fairly intuitive and effective way to maintain surface consistency. There are some caveats, however.
In order to produce a standard deviation greater than $0$, we must have a batch size greater than $1$.
Furthermore, environments that rarely show a particular region of a surface will rarely have $\lUC$ applied to that region because they're unlikely to show up twice in the same batch.
For this reason, $\lUC$ is most useful with large batch sizes.% because observing the same surface in multiple elements of a batch becomes more likely as the batch size increases.
But because video memory limits our batch sizes, first-person environments might be more of a challenge for $\lUC$ - particularly when exploring multiple rooms (a task we haven't done though - should I mention this?)

To michael: Should we mention these caveats in this section, or mention them in the texture realism loss section? That section addresses all of these issues...

% Also, you have to get lucky - two scene pictures have to show the same side of a cube for example. First person scenarios would be even worse.

\subsection{Texture Realism Loss}
\label{sec:texture_realism_loss}

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=250pt]{../images/texture_realism_ablation.pdf}
	\end{center}
	\caption{
		The neural texture looks more realistic with texture realism loss enabled. Note that the blocks show different letters because of different initializations and the symmetry of a cube - both solutions are valid texture assignments. 
	}
	\label{fig:texture_realism_ablation}
\end{figure}

We can improve the translation results by forcing the neural texture to look realistic, instead of allowing it to have arbitrary colors.
To do this, we add a texture realism loss $\lTR$.
$\lTR$ is an image similarity loss that makes $\pi$ look like its translation $\hat{p}$.
In other words, $\lTR$ makes the neural texture more realistic by making the neural texture projections more realistic.

So far, we've taken an image translator $T$ and added neural textures $\tau$ and a surface consistency constraint $\lUC$. 
If we wanted to, could stop here - those formulations are sufficient to produce consistent sim-to-real translations.
However, we can improve this further.

% Up until now we have not put any constraints on this neural texture. 
Without $\lTR$, $\tau$ can look wildly different each time you train it.
As seen in the top row of figure \ref{fig:texture_realism_ablation}, the neural texture looks very unrealistic - the colors are completely arbitrary.
After adding $\lTR$ though, $\pih \approx \ph$, and $\tau$ to represents the albedo of the surfaces in the scene.
The image similarity loss used in $\lTR$ is calculated the same way image similarity losses are calculated in section \ref{sec:image_translation}: L2 loss minus MSSSIM.

As it turns out, $\lTR$ is sufficient to ensure surface consistency without $\lUC$, though using both yields better results than either one alone. 
For an explanation of why this is the case, see the appendix \ref{sec:texture_realism_theory}.
There are some advantages to using $\lTR$ alone though: when just using $\lTR$ you can use batch sizes as small as $B=1$, allowing you to save video memory. 
Using $\lTR$ is also more numerically stable, as you no longer have to worry about whether the same surfaces are seen across multiple scenes in a batch.

% We introduce a new loss, called texture realism loss that fixes this: $\lTR$ is an image similarity loss that encourages $\pi \approx \ph$.
% This loss forces the neural texture to approximate the albedo of the surfaces in a scene.

% These images could be regularized so that they're the same every time.
% 
% To do this we put a penalty on the latent texture that makes it $\pi$ look like its translation $\hat{p}$. 
% 
% We use l2 and msssim losses to do this.
% 
% This helps improve the translations.
% 
% There are some advantages over unprojection consistency though: 
% smaller batch sizes (even a batch size of 1, which UC cant do),
% you don't have to get lucky (the more chaotic, or even the more zoomed in a dataset becomes the less likely there is to be any collisions)

\subsection{Color Loss}
ASK MICHAEL: I'd rather not mention this and run out the page limit...it's not very important, and not used in the robot experiments. It's only used in the experiment with the five items including the apple. Can we add it to the appendix instead?

@@

\clearpage

\section{Appendix} 
\label{sec:appendix}

	\subsection*{Unprojection Recovery Resolution Effect}

	\subsection*{Texture Realism Theory}
		\label{sec:texture_realism_theory}

		THIS SECTION IS UNPOLISHED, BUT CONTAINS  THE NESSECARY CONTENT.
		I'll delete this section if you don't think its important enough to keep.

		Texture realism loss equation:
		$L_{TR}=L_{TR_{l2}}+L_{TR_{msssim}}$
		where 
		$L_{TR_{msssim}} = -msssim(\pi,\hat{p})$ and
		$L_{TR_{l2}} = \frac{
			\sum\limits_{i=1}^B {
			%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
			\left( \pi-\hat{p}  \right)^2
				} 
		}
		{B}$

		(Give theoretical explation)
		$
		\ell_{UC} = 
		\
		\
		\mu\left(\sigma\left(u_{1\dots B,\ 1\dots L}\right)_{1 \dots L}\right) = 
		\sqrt{
		\frac{
			\sum\limits_{i=1}^B {
				%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
				\left( \hat{p}-\bar{p}  \right)^2
				} 
			}
			{B}
		}
		$
		Using $\bar{p}$ from figure \ref{fig:unprojection_consistency_loss}

		Becuase $\bar{p}$ is created from the average recovered texturs of $\hat{p}$, $\bar{p} \approx \hat{p}$.

		Becuse of texture realism loss, $\pi \approx \hat{p}$.

		Therefore, $\bar{p} \approx \hat{p} \approx \pi$

		% The l2 component of the texture realism loss equation:
		% $ \
		% \
		% \
		% \
		% \ \ell_{TR_{l2}}=\sum\limits_{i=1}^B {
		% 	%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
		% 	\left( \pi-\hat{p}  \right)^2
		% 	 } 
		% $

		Substituting one projected texture for another, aka $\pi$ for $\bar{p}$, we get 
		$ \
		\
		\
		\
		\ \ell_{TR_{l2}}
		\approx
		\frac{\sum\limits_{i=1}^B {
			%  \mu\left(u_{1\dots B,\ 1\dots L}\right)_{1...L}
			\left( \hat{p}-\pi  \right)^2
			}}{B} = \ell_{UC}^2
		$

>>>>

@@

\usepackage{corl_2022} % Use this for the initial submission.
% \usepackage[final]{corl_2022} % Uncomment for the camera-ready ``final'' version.
\usepackage{graphicx}
\usepackage{float}
\usepackage{amsfonts}

@@

\title{TRITON: Globally Consistent Sim-to-Real Transfer}

@@

@@

@@

\section{Data}
\label{sec:data}

@@

    \textbf{We need some general method description like below:}
    We learn neural texture corresponding to each object, where each texture is represented as a function mapping... 
    The idea is to learn object textures as a set of continuous functions from a set of (unlabeled) real images, and use them to generate realistic images.
    Given a 3D structure of the scene represented with a UVL image, the learned neural textures will translate them into a realistic image.
    When used for the robot learning, this enables learnable, realistic sim2real of robot training images.

@@

\subsection{Neural Texture}

@@

    The neural texture is a neural function not pixel grid.

	$B$ is the batch size. Given a batch of $B$ UVL Scenes $s_{1\dots B}$ with $L$ different label values, we project $L$ different neural textures $\tau_{1\dots L}$ onto that scene; obtaining projections $\tau_{1\dots B}$. This projection mechanism is illustrated in figure \ref{fig:uvl_explanation}.

@@

	Each neural texture $\tau_i$ is a function consisting of two components: a multi-layer perceptron $M_i$ and a static set of 256 random spatial frequencies $f_{1...256}$. 
	Given a two-dimensional $uv$ vector, the spatial frequencies are used to generate a high dimensional embedding used as input for the MLP.
	\makebox[\textwidth]
	{
		$
		\tau_i(uv) = 
		\
		M_i\left(\left[
			\sin f_1 uv, \ 
			\cos f_1 uv, \ 
			\sin f_2 uv, \ 
			\cos f_2 uv, \ 
			\dots \ 
			\sin f_{256} uv, \ 
			\cos f_{256} uv \ 
		\right]\right)
		$
	}
	where $uv$ and $f_i \in N_{ormal}(\mu=0,\ \sigma=10)$ are two dimensional vectors,
	 $f_i uv$ represents their dot product, 
	%  $rgb$ represents a 3d color vector,
	 and $M_i$ maps a $uv$ embedding to an $rgb$ color value:
	 $M_i: \mathbb{R}^{256} \rightarrow \mathbb{R}^3$.
	Note that there is nothing special about 256 or 10: these hyperparameter values just happen to work nicely.

@@

\subsection{Image Translation}
Our image translation module $T$ is based on the image translation module used in \citep{surgical_video_translation}, which is a variant of MUNIT \cite{munit}. Like in \citep{surgical_video_translation} but unlike MUNIT, however, the style code is fixed and noise is injected into the latent codes to prevent overfitting. During inference, our image translation module is completely determinstic.

$T$ is a bidirectional image translation module, and has two translation directions. $T_{A\rightarrow B}$ translates sim to real (aka domain A to domain B), and $T_{B\rightarrow A}$ translates real to sim (aka domain B to domain A). The simulation domain A (UVL maps concatenated to their texture projections) has 6 channels, whereas the real image domain B (photographs) has 3 channels.

Apart from these differences, the losses pertaining directly to image translation are the same as that in MUNIT.

Due to the way MUNIT works, we define
$T_{A\rightarrow B}\left(c\right) = dec_B\left(enc_A\left(c\right)\right)$
and 
$T_{B\rightarrow A}\left(p\right) = dec_A\left(enc_B\left(p\right)\right)$
where $enc_A, enc_B, dec_A, dec_B$ are encoders and decoders for domains A and B respectively, $p$ is a batch of real photos, and $c$ is a batch of simulated UVL+projection concatenations.

\subsection{Unprojection Consistency Loss}

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=400pt]{../images/unprojection_consistency_loss.pdf}
	\end{center}
	\caption{
		This is the unprojection consistency loss used in figure \ref{fig:main_diagram}.
		After calculating fake photos $\hat{p}_{1...B}$ from UVL scenes $s_{1...B}$ as seen in figure \ref{fig:main_diagram}, 
		we can unproject fake photos $\hat{p}_{1...B}$ back into texture space to obtain $u_{1...B,\ 1...L}$ .
		If the fake photos display consistent textures along their 3d surfaces, then the unprojected textures should be similar.
		We can measure this similarity by taking the per-channel standard deviation of these unprojections, aka $\sigma(u)$, then averaging the result.
		When the fake photos display similar surfaces, the unprojections will be similar making the unprojection consistency loss low.
		Additionally, we can also calculate the average unprojection, which is useful for visualization purposes (it will look blurry if $\hat{p}$ is inconsistent).
	}
	\label{fig:unprojection_consistency_loss}
\end{figure}

This benefits from large batch sizes, but that often makes you run out of vram.

Also, you have to get lucky - two scene pictures have to show the same side of a cube for example. First person scenarios would be even worse.

\subsection{Texture Realism Loss}

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=250pt]{../images/texture_realism_ablation.pdf}
	\end{center}
	\caption{
		The neural texture looks more realistic with texture realism loss enabled.
	}
	\label{fig:texture_realism_ablation}
\end{figure}

These images could be regularized so that they're the same every time.

To do this we put a penalty on the latent texture that makes it $\pi$ look like its translation $\hat{p}$.

We use L2 and MSSSIM losses to do this.

This helps improve the translations.

(Give theoretical explation)

There are some advantages over unprojection consistency though: 
smaller batch sizes (even a batch size of 1, which UC cant do),
you don't have to get lucky (the more chaotic, or even the more zoomed in a dataset becomes the less likely there is to be any collisions)

\subsection{Color Loss}

@@
